---
title: "Part 3: Top Parent Company Recipients"
output: github_document
---

In part 2, we looked into how grant & loan money was distributed amongst recipient companies. In part 3, we look more closely at the top 10 recipients.

## Load data

First, load the packages and read the data from part 1.

```{r, message=FALSE, warning = FALSE}
options(scipen = 999)
library(dplyr)
library(ggplot2) # for graphing
theme_set(theme_bw())

df <- read.csv('../data/formatted_loan_data_6_17.csv', header = TRUE)
```

## Analysis

##### What 10 parent companies received the most grant money? How about loans?

```{r}
### add values grouped by parent company
loan.grant.df <- df %>%
  filter(p_company != '') %>% # remove companies without parents
  group_by(award_type, p_company) %>% # specify variables to summarize by
  summarize(award_amount = sum(award_amount))

### extract top 10 values for printing and graphing
top10.df <- loan.grant.df %>%
  ungroup() %>%
  group_by(award_type) %>%
  arrange(desc(award_amount)) %>% # arrange from highest to lowest
  slice(1:10)

print(top10.df)
```

Since we're looking at the highest values, we may want to convert our dollar units into millions of dollars for graphing.

```{r}
top10.df$award_amount_mil <- top10.df$award_amount / 1000000

p1 <- ggplot(top10.df %>% group_by(p_company)) +
  geom_bar(aes(x = p_company, y = award_amount_mil), stat = 'identity') +
  facet_grid(award_type ~ .)

p1
```

This visualization shows that: 

* 4 of the top 10 loan recipients were also in the top 10 for money received in grants
* a stacked bar graph may be a better option

So, let's try a stacked bar graph.

```{r, fig.width=14}
top10.df$company_title <- gsub(' |-', '\n', top10.df$p_company) # wrap x-axis labels so they don't overlap

p2 <- ggplot(top10.df) +
  geom_bar(aes(x = company_title, y = award_amount_mil, fill = award_type), stat = 'identity') +
  labs(
    title = 'Total COVID-19 Loans and Grants by Parent Company',
    x = 'Company',
    y = 'Dollars (in billions)'
  )
p2
```

This helps us better visualize the data that is mixed between loans and grants with some overlapping companies. I want to note a couple of observations:

* All of the recipients that fit our criteria were either in the healthcare or airline industries
* American, Delta, Southwest, and United Airlines received a massive mix of loans and grants

Let's look a bit more into the top recipients.

##### Of the top 10 companies, what percentage of grants did they receive relative to the whole?

```{r}
all.grants <- sum(df$award_amount[df$award_type == 'grant'])
top.grants <- sum(top10.df$award_amount[top10.df$award_type == 'grant'])

perc.grants <- (top.grants / all.grants) * 100
perc.grants <- round(perc.grants, 2)

print(perc.grants)
```

It turns out that 10 parent companies received about `r perc.grants`% of all grant money. 

##### Of the top 10 companies, what percentage of loans did they receive relative to the whole?

```{r}
all.loans <- sum(df$award_amount[df$award_type == 'loan'])
top.loans <- sum(top10.df$award_amount[top10.df$award_type == 'loan'])

perc.loans <- (top.loans / all.loans) * 100
perc.loans <- round(perc.loans, 2)

print(perc.loans)
```

The top 10 parent companies received about `r perc.loans`% of all grant money. 
