---
title: "COVID-19 Facts and Figures"
output: github_document
---

```{r, include = FALSE, warning = FALSE, message = FALSE}
options(scipen = 999)
library(dplyr)
library(ggplot2)
library(ggthemr) # set theme for graphs
library(kableExtra)
source('../plot_themes/bar_graph_theme.R')
source('../functions/add_thousands_comma.R')
```

```{r, include = FALSE, warning = FALSE, message = FALSE}
##################
##### Part 2 #####
##################

### Graphs
pt2.grant.p1 <- readRDS('../2_Overall_Distributions/r_objects/p_lower_90_grant.rds') # histograms of lower 90%
pt2.loan.p1 <- readRDS('../2_Overall_Distributions/r_objects/p_lower_90_loan.rds')

pt2.grant.p2 <- readRDS('../2_Overall_Distributions/r_objects/p_upper_10_grant.rds') # histograms of upper 0.5%-10%
pt2.loan.p2 <- readRDS('../2_Overall_Distributions/r_objects/p_upper_10_loan.rds')

pt2.grant.p3 <- readRDS('../2_Overall_Distributions/r_objects/p_upper_05_grant.rds') # histograms of upper 0.5%
pt2.loan.p3 <- readRDS('../2_Overall_Distributions/r_objects/p_upper_05_loan.rds')

### Other
pt2.grant.perc1 <- readRDS('../2_Overall_Distributions/r_objects/grant_500k_1m_perc.rds') # percent recipients of between 500k & 1m 
pt2.loan.perc1 <- readRDS('../2_Overall_Distributions/r_objects/loan_500k_1m_perc.rds')

pt2.grant.df1 <- readRDS('../2_Overall_Distributions/r_objects/grant_400m_plus_comps.rds') # names of recipients of $400m or more
pt2.loan.df1 <- readRDS('../2_Overall_Distributions/r_objects/loan_400m_plus_comps.rds')
```

```{r, include = FALSE}
### data munging
pt2.grant.df1$`Grant Amount (millions)` <- add_thousands_comma(pt2.grant.df1$`Grant Amount (millions)`)
pt2.grant.df1$`Grant Amount (millions)` <- paste('$', pt2.grant.df1$`Grant Amount (millions)`)

pt2.loan.df1$`Loan Amount (millions)` <- add_thousands_comma(pt2.loan.df1$`Loan Amount (millions)`)
pt2.loan.df1$`Loan Amount (millions)` <- paste('$', pt2.loan.df1$`Loan Amount (millions)`)
```

```{r, include = FALSE, warning = FALSE, message = FALSE}
##################
##### Part 3 #####
##################

### Graphs
pt3.p1 <- readRDS('../3_Top_Parent_Companies/r_objects/top_10_loans_grants_graph.rds') # bar graph of top 10 parent company recipients of loans and grants

### Other
pt3.df1 <- readRDS('../3_Top_Parent_Companies/r_objects/top_10_loans_grants.rds') # data frame of top 10 parent company recipients of loans and grants

pt3.perc.grants <- readRDS('../3_Top_Parent_Companies/r_objects/top_10_grants_received_perc.rds') # percentage of money in grants received by top 10 relative to all allocated money
pt3.perc.loans <- readRDS('../3_Top_Parent_Companies/r_objects/top_10_loans_received_perc.rds') # percentage of money in loans received by top 10 relative to all allocated money
```

```{r, include = FALSE}
### data munging
pt3.grant.df1 <- pt3.df1 %>% 
  ungroup() %>%
  filter(award_type == 'grant') %>%
  select(`Parent Company` = p_company, award_amount_mil) %>%
  mutate(`Grant Amount (billions)` = round((award_amount_mil / 1000), 2)) %>%
  mutate(`Grant Amount (billions)` = format(`Grant Amount (billions)`, scientific = FALSE, digits = 2)) %>%
  select(-award_amount_mil)

pt3.loan.df1 <- pt3.df1 %>% 
  filter(award_type == 'loan') %>%
  ungroup() %>%
  select(`Parent Company` = p_company, award_amount_mil) %>%
  mutate(`Loan Amount (billions)` = round((award_amount_mil / 1000), 2)) %>%
  mutate(`Loan Amount (billions)` = format(`Loan Amount (billions)`, scientific = FALSE, digits = 2)) %>%
  select(-award_amount_mil)

### for tables
pt3.grant.df1$`Grant Amount (billions)` <- paste('$', pt3.grant.df1$`Grant Amount (billions)`)
pt3.loan.df1$`Loan Amount (billions)` <- paste('$', pt3.loan.df1$`Loan Amount (billions)`)

### for in-line text
pt3.total.grant.money <- paste0('$', round(sum(pt3.df1$award_amount[pt3.df1$award_type == 'grant']) / 1000000000, 1))
pt3.total.loan.money <- paste0('$', round(sum(pt3.df1$award_amount[pt3.df1$award_type == 'loan']) / 1000000000, 1))
```

# Facts

* <B>`r pt2.grant.perc1`%</B> of grant recipients were given between \$500k and \$1m
* <B>`r pt2.loan.perc1`%</B> of loan recipients were given between \$500k and \$1m
* <B>`r nrow(pt2.grant.df1)`</B> grant recipients were given \$400m or more
* <B>`r nrow(pt2.loan.df1)`</B> loan recipients were given \$400m or more

* The top 10 parent company grant recipients were given around <B>`r pt3.total.grant.money` billion dollars</B> - about <B>`r pt3.perc.grants`%</B> of the distributed grant funds
* The top 10 parent company loan recipients were given around <B>`r pt3.total.loan.money` billion dollars</B> - about <B>`r pt3.perc.loans`%</B> of the distributed loan funds

# Tables

```{r, echo = FALSE, fig.show="hold", out.width="50%", warning = FALSE, message = FALSE}
kable(pt2.grant.df1, caption = 'Companies Receiving COVID-19 Grants of over $400m')
```

```{r, echo = FALSE, fig.show="hold", out.width="50%", warning = FALSE, message = FALSE}
kable(pt2.loan.df1, caption = 'Companies Receiving COVID-19 Loans of over $400m')
```

```{r, echo = FALSE, fig.show="hold", out.width="50%", warning = FALSE, message = FALSE}
kable(pt3.grant.df1, caption = 'Top 10 Parent Companies Receiving COVID-19 Grants')
```

```{r, echo = FALSE, fig.show="hold", out.width="50%", warning = FALSE, message = FALSE}
kable(pt3.loan.df1, caption = 'Top 10 Parent Companies Receiving COVID-19 Loans')
```

# Graphs

## Money Distributions

```{r, echo = FALSE, warning = FALSE, message = FALSE}
pt2.grant.p1; pt2.loan.p1
```

```{r, echo = FALSE, warning = FALSE, message = FALSE}
pt2.grant.p2; pt2.loan.p2
```

```{r, echo = FALSE, warning = FALSE, message = FALSE}
pt2.grant.p3; pt2.loan.p3
```

## Top 10 Companies

```{r, fig.width=14.5, echo = FALSE, warning = FALSE, message = FALSE}
pt3.p1
```
