---
title: "Results: COVID-19 Loan Distributions"
output: github_document
---

All information was gathered using data from https://data.covidstimuluswatch.org/prog.php?&detail=export_csv . Some recipient names are cut off in the file.

```{r, include = FALSE, warning = FALSE, message = FALSE}
library(ggplot2)
library(ggthemr) # set theme for graphs
library(kableExtra)
source('../plot_themes/bar_graph_theme.R')
source('../functions/add_thousands_comma.R')

### Graphs
grant.p1 <- readRDS('../2_Overall_Distributions/r_objects/p_lower_90_grant.rds') # histograms of lower 90%
loan.p1 <- readRDS('../2_Overall_Distributions/r_objects/p_lower_90_loan.rds')

grant.p2 <- readRDS('../2_Overall_Distributions/r_objects/p_upper_10_grant.rds') # histograms of upper 0.5%-10%
loan.p2 <- readRDS('../2_Overall_Distributions/r_objects/p_upper_10_loan.rds')

grant.p3 <- readRDS('../2_Overall_Distributions/r_objects/p_upper_05_grant.rds') # histograms of upper 0.5%
loan.p3 <- readRDS('../2_Overall_Distributions/r_objects/p_upper_05_loan.rds')

### Other
grant.perc1 <- readRDS('../2_Overall_Distributions/r_objects/grant_500k_1m_perc.rds') # percent recipients of between 500k & 1m 
loan.perc1 <- readRDS('../2_Overall_Distributions/r_objects/loan_500k_1m_perc.rds')

grant.df1 <- readRDS('../2_Overall_Distributions/r_objects/grant_400m_plus_comps.rds') # names of recipients of $400m or more
loan.df1 <- readRDS('../2_Overall_Distributions/r_objects/loan_400m_plus_comps.rds')
```

```{r, include = FALSE}
### data munging
grant.df1$`Grant Amount (millions)` <- add_thousands_comma(grant.df1$`Grant Amount (millions)`)
grant.df1$`Grant Amount (millions)` <- paste('$', grant.df1$`Grant Amount (millions)`)

loan.df1$`Loan Amount (millions)` <- add_thousands_comma(loan.df1$`Loan Amount (millions)`)
loan.df1$`Loan Amount (millions)` <- paste('$', loan.df1$`Loan Amount (millions)`)
```

## Facts

* <B>`r grant.perc1`%</B> of grant recipients were given between \$500k and \$1m
* <B>`r loan.perc1`%</B> of loan recipients were given between \$500k and \$1m
* <B>`r nrow(grant.df1)`</B> grant recipients were given \$400m or more
* <B>`r nrow(loan.df1)`</B> loan recipients were given \$400m or more

* The top 10 parent company grant recipients were given around <B>`r total.grant.money` billion dollars</B> - about <B>`r perc.grants`%</B> of the distributed grant funds
* The top 10 parent company loan recipients were given around <B>`r total.loan.money` billion dollars</B> - about <B>`r perc.loans`%</B> of the distributed loan funds

## Tables

```{r, echo = FALSE, fig.show="hold", out.width="50%", warning = FALSE, message = FALSE}
kable(grant.df1, caption = 'Companies Receiving COVID-19 Grants of over $400m')
```

```{r, echo = FALSE, fig.show="hold", out.width="50%", warning = FALSE, message = FALSE}
kable(loan.df1, caption = 'Companies Receiving COVID-19 Loans of over $400m')
```

## Graphs

The below graphs show how COVID-19 grants & loans were distributed amongst the lower 90% of recipients

```{r, echo = FALSE, warning = FALSE, message = FALSE}
grant.p1; loan.p1
```

The below graphs show how COVID-19 grants & loans were distributed amongst the top 10% through top 0.5% of recipients

```{r, echo = FALSE, warning = FALSE, message = FALSE}
grant.p2; loan.p2
```

The below graphs show how COVID-19 grants & loans were distributed amongst the top 0.5% of recipients

```{r, echo = FALSE, warning = FALSE, message = FALSE}
grant.p3; loan.p3
```
