---
title: "Part 2: COVID-19 Loan Distributions"
output: github_document
---

For part 2, we are going to look at how loans and grants are distributed amongst companies and parent companies. We will get into dividing the companies based on research questions in later pages.

## Load data

First, load the packages and read the data from part 1.

```{r, message=FALSE, warning = FALSE}
options(scipen = 999) # to avoid exponent symbol in number representations
library(dplyr)
library(ggplot2) # for graphing
theme_set(theme_bw()) # because I refuse to use the default ggplot2 theme

df <- read.csv('../data/formatted_loan_data_6_17.csv', header = TRUE)
```

## Analysis

##### How is grant money distributed amongst recipients?

Your best 2 options for visualization of distributions would be a histogram or a density plot. We will go with the former.

```{r, message = FALSE, warning = FALSE}
df.grant <- df %>% 
  filter(award_type == 'grant' & grant_amount > 0) %>% # extract grants only
  mutate(grant_amount_mil = grant_amount / 1000000) # convert to millions

ggplot(df.grant) +
  geom_histogram(aes(grant_amount_mil), bins = 30)
```

As we will see in part 3, there are some outliers that ruin the scaling of this graph. Since we look at this later, we can remove them for the purpose of this page. Let's cut off, say, the top 10%.

```{r, message = FALSE, warning = FALSE}
df.grant.90 <- df.grant %>% filter(grant_amount < (quantile(df.grant$grant_amount, .90)))

grant.p2 <- ggplot(df.grant.90) +
  geom_histogram(aes(grant_amount_mil), bins = 30) +
  labs(
    title = 'Companies Receiving COVID-19 Grants',
    x = 'Grant Amount\n(millions of dollars)',
    y = '# of Companies'
  )
grant.p2
```

That's a lot better. It looks like a pretty solid number of the companies received loans between \$500k and \$1m. How many of them is that compared to the whole?

```{r, message = FALSE, warning = FALSE}
num.between <- df.grant %>%
  filter(grant_amount >= 500000 & grant_amount <= 1000000) %>%
  distinct(recipient_company) %>% # remove companies that received multiple grants
  nrow()

grant.perc.500k.1m <- (num.between / (nrow(df.grant %>% distinct(recipient_company)))) * 100
grant.perc.500k.1m <- round(grant.perc.500k.1m, 2)

print(grant.perc.500k.1m)
```

So about `r grant.perc.500k.1m`% of companies received a grant between \$500,000 and \$1,000,000.

##### How is loan money distributed amongst recipients?

Now we want to look at loans through the same process.

```{r}
df.loan <- df %>% 
  filter(award_type == 'loan' & face_loan_amount > 0) %>%
  mutate(face_loan_amount_mil = face_loan_amount / 1000000)
```

We remove the top 10% again to be consistent.

```{r, message = FALSE, warning = FALSE}
df.loan.90 <- df.loan %>% filter(face_loan_amount < (quantile(df.loan$face_loan_amount, .90)))

loan.p1 <- ggplot(df.loan.90) +
  geom_histogram(aes(face_loan_amount_mil), bins = 30) +
  labs(
    title = 'Companies Receiving COVID-19 Loans',
    x = 'Loan Amount\n(millions of dollars)',
    y = '# of Companies'
  )

loan.p1
```

The loans also seem to be most commonly between \$500k and \$1m. How many of them is that compared to the whole?

```{r, message = FALSE, warning = FALSE}
num.between <- df.loan %>%
  filter(face_loan_amount >= 500000 & face_loan_amount <= 1000000) %>%
  distinct(recipient_company) %>% # remove companies that received multiple grants
  nrow()

loan.perc.500k.1m <- (num.between / (nrow(df.loan %>% distinct(recipient_company)))) * 100
loan.perc.500k.1m <- round(loan.perc.500k.1m, 2)
print(loan.perc.500k.1m)
```

So about `r loan.perc.500k.1m`% of companies received a loan between \$500,000 and \$1,000,000. Extremely close to the percentage we see with grants.

## Top 10%

We looked at how the lower 90% is distributed. But how do companies that were in the top 10% look?

##### How is grant money distributed amongst the upper 10%?

```{r, message = FALSE, warning = FALSE}
df.grant.10 <- df.grant %>% filter(grant_amount > (quantile(df.grant$grant_amount, .90)))

grant.p3 <- ggplot(df.grant.10) +
  geom_histogram(aes(grant_amount_mil), bins = 30)

grant.p3
```

Like the lower 90%, this isn't exactly an ideal graph scaling. For graphing, we'll have to split these up. One way to find a cutoff value is to find what values are influencing the weights of ggplot2's scaling.

```{r}
quants <- sapply(seq(0, 1, 0.05), function(x) { quantile(df.grant.10$grant_amount, x) })
print(quants)
```

Just by eyeballing, the largest gap in quantiles is pretty clearly between 95% and 100%.

```{r}
diff(quants)
```

This shows the difference in dollars between x percentile and x-1 percentile cutoffs. Seems to support the above observation, so we'll just look at the lower 95% (of the overall top 10%). This isn't to say that the top 5% is not of interest - it just isn't part of the current goal.

```{r, message = FALSE, warning = FALSE}
df.grant.10 <- df.grant.10 %>% filter(grant_amount < (quantile(df.grant.10$grant_amount, .95)))

grant.p4 <- ggplot(df.grant.10) +
  geom_histogram(aes(grant_amount_mil), bins = 30) +
  labs(
    title = 'Top 10% Companies Receiving COVID-19 Grants',
    caption = '*with top 5% removed for graph scaling',
    x = 'Grant Amount\n(millions of dollars)',
    y = '# of Companies'
  )

grant.p4
```

## Now rinse and repeat

##### How is loan money distributed amongst the upper 10%?

```{r, message = FALSE, warning = FALSE}
df.loan.10 <- df.loan %>% filter(face_loan_amount > (quantile(df.loan$face_loan_amount, .90)))
df.loan.10.2 <- df.loan.10 %>% filter(face_loan_amount < (quantile(df.loan.10$face_loan_amount, .95)))

loan.p2 <- ggplot(df.loan.10.2) +
  geom_histogram(aes(face_loan_amount_mil), bins = 30) +
  labs(
    title = 'Top 10% Companies Receiving COVID-19 Loans',
    caption = '*with top 5% removed for graph scaling',
    x = 'Loan Amount\n(millions of dollars)',
    y = '# of Companies'
  )

loan.p2
```

## Top 0.5%

We have looked at the lower 90% of companies and the top 10% of companies with the top 5% of that removed. So, in terms of overall percentages, our last stop is looking at the top 0.5%. 

##### How is grant money distributed amongst the top 0.5% of recipients?

```{r, message = FALSE, warning = FALSE}
df.grant.05 <- df.grant %>% filter(grant_amount > (quantile(df.grant$grant_amount, .995)))

grant.p5 <- ggplot(df.grant.05) +
  geom_histogram(aes(grant_amount_mil), bins = 30) +
  labs(
    title = 'Top 0.5% Companies Receiving COVID-19 Grants',
    x = 'Grant Amount\n(millions of dollars)',
    y = '# of Companies'
  )

grant.p5
```

It looks like our outliers received $400 million or more in grants. What companies are these?

```{r}
fohund.milplus <- df.grant %>%
  filter(grant_amount_mil >= 400) %>%
  select(recipient_company, grant_amount_mil)

print(fohund.milplus)

fohund.print <- paste(fohund.milplus$recipient_company, collapse = ', ') # for dynamic markdown content
```

`r nrow(fohund.milplus)` companies received \$400 million or more in grants. These companies are `r fohund.print`.

##### How is loan money distributed amongst the top 0.5% of recipients?

```{r, message = FALSE, warning = FALSE}
df.loan.05 <- df.loan %>% filter(face_loan_amount > (quantile(df.loan$face_loan_amount, .995)))
 
loan.p3 <- ggplot(df.loan.05) +
  geom_histogram(aes(face_loan_amount_mil), bins = 30) +
  labs(
    title = 'Top 0.5% Companies Receiving COVID-19 Loans',
    x = 'Loan Amount\n(millions of dollars)',
    y = '# of Companies'
  )

loan.p3
```

Like grants, it looks like our outliers received $400 million or more in loans. What companies are these?

```{r}
fohund.milplus2 <- df.loan %>%
  filter(face_loan_amount_mil >= 400) %>%
  mutate(recipient_company = as.character(recipient_company))

fohund.milplus2 <- fohund.milplus2 %>%
  arrange(desc(face_loan_amount_mil)) %>%
  select(recipient_company, face_loan_amount_mil) 
print(fohund.milplus2)

fohund.print2 <- paste(fohund.milplus2$recipient_company, collapse = ', ') 
```

`r nrow(fohund.milplus2)` companies received \$400 million or more in loans. These companies are `r fohund.print2`.
